---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(ggplot2)
library(tsibble)
library(tsibbledata)
library(lubridate)
```

>data from end of 2016 - end of 2021
>KPIs: staffed beds available
       percentage occupancy
>metrics: vs time - average of whole dataset
                  - by location (hb) - rural vs urban areas
                  - by specialty name - cardiology, mental health, emergency
          
          > forecast based on pre covid data and compare to actual with all data      
          
```{r}
area_codes <- read_csv("raw_data/general/hb14_hb19.csv") %>% 
  janitor::clean_names()

beds_treatment_specialty <- read_csv("raw_data/non_covid/beds_by_nhs_board_of_treatment_and_specialty.csv") %>% 
  janitor::clean_names()

beds_treatment_specialty <- beds_treatment_specialty %>% 
  mutate(percentage_occupancy = coalesce(percentage_occupancy, median(percentage_occupancy, na.rm = TRUE)))
```


## time series of avg occupancy by location 
```{r}
beds_available <- beds_treatment_specialty %>% 
  mutate(quarter = yq(quarter), month = month(quarter, label = TRUE, abbr = TRUE),
         year = year(quarter)) %>%
  select(quarter, month, year, hb, specialty_name, average_available_staffed_beds, percentage_occupancy) %>% 
  distinct()

beds_available <- left_join(beds_available, area_codes, "hb") %>% 
  select(-c(hb_date_enacted, hb_date_archived, country)) %>% mutate(urban_area = if_else(hb_name %in% c("NHS Greater Glasgow and Clyde", "NHS Lanarkshire", "NHS Lothian"), "high population", "low population")) #based on greater than 500,000

beds_available <- tsibble(beds_available, index = "quarter", key = c(hb, month, year, specialty_name, average_available_staffed_beds, percentage_occupancy, hb_name, urban_area))

beds_available %>% 
  group_by(hb_name) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy)) %>% 
  filter(hb_name == "NHS Ayrshire and Arran") %>% #sub out for desired area
  ggplot()+
  geom_line(aes(x = quarter, y = avg_occupancy))+
  labs(title = "Avg. Beds Available by Region",
       x = "Date",
       y = "Percentage Occupancy")
```
##show all locations as percentage
```{r}
beds_available %>% 
  filter(!is.na(hb_name)) %>% 
  group_by(hb_name) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy)) %>% 
  ggplot()+
  geom_line(aes(x = quarter, y = avg_occupancy, colour = hb_name))+
  labs(title = "Avg. Beds Available by Region",
       x = "Date",
       y = "Percentage Occupancy")

unique(beds_available$hb_name)
```


## time series of avg occupancy by urban population size
```{r}
beds_available %>% 
  group_by(urban_area) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy)) %>% 
  filter(urban_area == "low population") %>% #sub out for desired 
  ggplot()+
  geom_line(aes(x = quarter, y = avg_occupancy))+
  labs(title = "Avg. Beds Available by Population Status",
       x = "Date",
       y = "Percentage Occupancy")
```
## show all population status' as percentage
```{r}
beds_available %>% 
  group_by(urban_area) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy))%>%  ggplot()+
  geom_line(aes(x = quarter, y = avg_occupancy, colour = urban_area))+
  labs(title = "Avg. Beds Available by Population Status",
       x = "Date",
       y = "Percentage Occupancy")
```




## time series of avg occupancy overall
```{r}
beds_available %>% 
  index_by(quarter) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy)) %>% 
  ggplot()+
  geom_line(aes(x = quarter, y = avg_occupancy))+
  labs(title = "Avg. Beds Available",
       x = "Date",
       y = "Percentage Occupancy")
```
## time series of avg occupancy by specialty - emergency grouping - cardiology - mental health grouping
```{r}
beds_available %>% 
  group_by(specialty_name) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy)) %>% 
  filter(specialty_name == "Mental Health Grouping" | specialty_name == "Emergency Grouping" | specialty_name == "Cardiology") %>% #sub out for desired area
  ggplot()+
  geom_line(aes(x = quarter, y = avg_occupancy, colour = specialty_name))+
  labs(title = "Avg. Beds Available by Specialty",
       x = "Date",
       y = "Percentage Occupancy")


unique(beds_available$sp)
```

## Seasonality of occupancy
```{r}
beds_treatment_specialty %>% 
  mutate(date = yq(quarter), month = month(date, label = TRUE, abbr = TRUE)) %>% 
  group_by(month) %>% 
  summarise(avg_beds = mean(percentage_occupancy)) %>% 
  ggplot(aes(x = factor(month, levels = c("Jul", "Oct", "Jan", "Apr")), y = avg_beds))+
  geom_point()+
  geom_line(group = 1)+
  labs(title = "Percentage Occupancy Seasonality",
       x = "Month",
       y = "Perecentage Occupncy")

beds_treatment_specialty %>% 
  mutate(date = yq(quarter), month = month(date, label = TRUE, abbr = TRUE),
         year = year(date)) %>% 
  filter(year <= 2019) %>% 
  group_by(month) %>% 
  summarise(avg_beds = mean(percentage_occupancy)) %>% 
  ggplot(aes(x = factor(month, levels = c("Jul", "Oct", "Jan", "Apr")), y = avg_beds))+
  geom_point()+
  geom_line(group = 1)+
  labs(title = "Percentage Occupancy Seasonality",
       x = "Month",
       y = "Perecentage Occupncy")

beds_treatment_specialty %>% 
  mutate(date = yq(quarter), month = month(date, label = TRUE, abbr = TRUE),
         year = year(date)) %>% 
  filter(year > 2019) %>% 
  group_by(month) %>% 
  summarise(avg_beds = mean(percentage_occupancy)) %>% 
  ggplot(aes(x = factor(month, levels = c("Jul", "Oct", "Jan", "Apr")), y = avg_beds))+
  geom_point()+
  geom_line(group = 1)+
  labs(title = "Percentage Occupancy Seasonality",
       x = "Month",
       y = "Perecentage Occupncy")
```
```{r}
beds_treatment_specialty <- left_join(beds_treatment_specialty, area_codes, "hb") %>% 
  select(-c(hb_date_enacted, hb_date_archived, country)) %>% mutate(urban_area = if_else(hb_name %in% c("NHS Greater Glasgow and Clyde", "NHS Lanarkshire", "NHS Lothian"), "high population", "low population"))

beds_treatment_specialty %>% #pre covid dates, all health baords
  mutate(date = yq(quarter), month = month(date, label = TRUE, abbr = TRUE),
         year = year(date)) %>% 
  filter(year <= 2019) %>% 
  group_by(month) %>% 
  summarise(avg_beds = mean(percentage_occupancy)) %>% 
  ggplot(aes(x = factor(month, levels = c("Jul", "Oct", "Jan", "Apr")), y = avg_beds))+
  geom_point()+
  geom_line(group = 1)+
  labs(title = "Percentage Occupancy Seasonality",
       x = "Month",
       y = "Perecentage Occupncy")

beds_treatment_specialty %>% 
  mutate(date = yq(quarter), month = month(date, label = TRUE, abbr = TRUE),
         year = year(date)) %>% 
  filter(year > 2019,
         hb_name == "NHS Lothian") %>% #filter based on map input
  group_by(month) %>% 
  summarise(avg_occupancy = mean(percentage_occupancy)) %>%
  ggplot(aes(x = factor(month, levels = c("Jul", "Oct", "Jan", "Apr")), y = avg_occupancy))+
  geom_point()+
  geom_line(group = 1)+
  labs(title = "Percentage Occupancy Seasonality",
       x = "Month",
       y = "Perecentage Occupncy")

```

##POTENTIAL PLOT FOR DASHBOARD
```{r}
df1 <- beds_treatment_specialty %>% #pre covid dates, all health baords
  separate(quarter, c("year", "quarter"), sep = "Q") %>% 
  filter(year <= 2019, hb_name == "NHS Lothian") %>% # hb selected by user input
  group_by(quarter) %>% 
  summarise(avg_occupancy = sum(total_occupied_beddays)/sum(all_staffed_beddays))


df2 <- beds_treatment_specialty %>% 
  separate(quarter, c("year", "quarter"), sep = "Q") %>% 
  filter(year > 2019, hb_name == "NHS Lothian") %>% #post covid, filter based on map input
  group_by(year, quarter) %>% 
  summarise(avg_occupancy = sum(total_occupied_beddays)/sum(all_staffed_beddays))


df2 %>% 
  ggplot(aes(x = quarter, y = avg_occupancy, colour = factor(year), group = year))+
  geom_point() +
  geom_line() +
  geom_line(data = df1, aes(x = quarter, y = avg_occupancy, group = 1, col = "Baseline")) +
  geom_point(data = df1, aes(x = quarter, y = avg_occupancy, group = 1, col = "Baseline")) +
  scale_color_manual(name = "Year", values = c("Baseline" = "darkblue",
                                               "2020" = "red",
                                               "2021" = "green")) +
  labs(title = "Avg Occupancy Bed Occupancy vs Baseline",
       x = "Quarter",
       y = "Avg Occupancy Rate")
```
```{r}
beds_treatment_specialty %>%
  filter(hb == "S08000015") %>% 
  separate(quarter, c("year", "quarter"), sep = "Q") %>% 
  group_by(quarter) %>% 
  summarise(sum = sum(total_occupied_beddays)/sum(all_staffed_beddays))

beds_treatment_specialty %>%
  filter(hb == "S08000015") %>% 
  separate(quarter, c("year", "quarter"), sep = "Q") %>% 
  group_by(quarter) %>% 
  summarise(sum = mean(percentage_occupancy))
```

