---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(tsibble)
library(here)
library(reshape2)
library(ggplot2)
library(infer)
library(scales)
```

```{r}
covid_age_sex_hb <- read_csv(here("raw_data/covid_data/hospital_admissions_hb_agesex_20220302.csv"))
covid_simd_hb <-read_csv(here("raw_data/covid_data/hospital_admissions_hb_simd_20220302.csv"))
covid_specialty_hb <- read_csv(here("raw_data/covid_data/hospital_admissions_hb_specialty_20220302.csv"))
hb <- read_csv(here("raw_data/general/hb14_hb19.csv")) %>% janitor::clean_names()
```

------------   cleaning ---------
I will clean 3 hb related data set and join them together at the end 

2.3. Data transformations including data cleaning
### Create date from the week_ending columns and select
### the useful columns that we will be looking at. 
```{r}
covid_age_sex_short <- covid_age_sex_hb %>% janitor::clean_names() %>% 
  select(c(week_ending, hb, age_group, sex, admission_type, number_admissions, average20182019, percent_variation)) %>% 
  mutate(week_ending = as.Date(ymd(week_ending)))
  
```

```{r}
covid_simd_short <- covid_simd_hb %>% janitor::clean_names() %>% 
  select(c(week_ending, hb, simd_quintile, admission_type, number_admissions, average20182019, percent_variation)) %>% 
  mutate(week_ending = as.Date(ymd(week_ending)))
```


```{r}
covid_specialty_short <- covid_specialty_hb %>% janitor::clean_names() %>% 
  select(c(week_ending, hb, specialty, admission_type, number_admissions, average20182019, percent_variation)) %>% 
  mutate(week_ending = as.Date(ymd(week_ending)))
```

```{r}
covid_hb <- covid_age_sex_short %>% full_join(covid_simd_short)%>% full_join(covid_specialty_short)
```

2.5. Exploratory data analysis
emergency vs planned admission type overall count 
```{r}
covid_hb %>% 
  filter(admission_type %in% c("Emergency","Planned")) %>% 
  drop_na(number_admissions) %>% 
  ggplot(aes(week_ending, number_admissions))+
  geom_line(aes(col = admission_type), group = 1)+
  facet_wrap( ~ admission_type, ncol = 1, scales = "free_y")+
  labs(title = "draft- admission_type compare (numbers/counts)")
```
```{r}
covid_hb %>% 
  filter(admission_type %in% c("Emergency","Planned")) %>% 
  drop_na(number_admissions) %>% 
  ggplot(aes(week_ending, number_admissions))+
  geom_col(aes(fill = admission_type), position = "fill")+
  labs(title = "draft- admission_type compare (proportion)",
       subtitle = "...very ewww, I don't like this graph")
```

-----emergency---
tidyed a bit (all group by week, mean of all hb)
```{r}
covid_emergency <- covid_hb %>% 
  group_by(week_ending) %>% 
  filter(admission_type == "Emergency",
         age_group == "All ages",
         sex == "All",
         specialty == "All") %>% 
  summarise(mean_this_week = mean(number_admissions),
            mean_1819 = mean(average20182019),
            var = mean(percent_variation))
```




```{r}



data_mod_emergency <- melt(covid_emergency,
                  measure.vars=c('mean_this_week', 'mean_1819'))
  

ggplot(data_mod_emergency) +
geom_boxplot(aes(y=value, color=variable))+
  labs(title = "emergency admission all age, all sex, all specialty, mean of all hb")
```


 
2.6. Statistical analyses to identify patterns, trends and relationships in the dataset
### Over time trend of percent variation for emergency admissions
```{r}
covid_emergency %>% 
  ggplot(aes(week_ending, var)) +
    geom_col(fill = case_when(covid_emergency$var >= 0 ~"red",
                              TRUE ~ "seagreen"))+
  labs(title = "emergency percent variation comparing with 2018/2019 ")
```

-------planned -------
repeated as emergency 
2.6. Statistical analyses to identify patterns, trends and relationships in the dataset
```{r}
covid_planned <- covid_hb %>% 
  group_by(week_ending) %>% 
  filter(admission_type == "Planned",
         age_group == "All ages",
         sex == "All",
         specialty == "All") %>% 
  summarise(mean_this_week = mean(number_admissions),
            mean_1819 = mean(average20182019),
            var = mean(percent_variation))
```
```{r}
data_mod_planned <- melt(covid_planned,
                  measure.vars=c('mean_this_week', 'mean_1819'))
  

ggplot(data_mod_planned) +
geom_boxplot(aes(y=value, color=variable))+
  labs(title = "planned admission, mean of all hb",
       subtitle = "all age, all sex, all specialty ")
```

``` 
2.7. Interpreting the results of statistical analysis
From this box plot we can see that on average, planned admssoin ini 2020(i.e. Mean this week)
is a lot lower than the mean of this figure in 2018 and 2019. This suggested that during covid
people didn't have as many planned admissions to the hospitals as before covid.
```

```{r}
covid_planned %>% 
  ggplot(aes(week_ending, var)) +
    geom_col(fill = case_when(covid_planned$var >= 0 ~"red",
                              TRUE ~ "seagreen"))+
  labs(title = "draft-planned percent variation comparing with 2018/2019 ",
       subtitle = "why! i don' understand, i must have done something wrong",
       fill = "fix it later")
```
 
 
 ----I will add few more cols for analysis later -----
 
```{r}
covid_full <- hb %>% janitor::clean_names() %>% 
  select(1,2) %>% 
right_join(covid_hb,by = "hb") 

covid_full <- covid_full %>% mutate(
  year = year(week_ending),
  month = month(week_ending,label = TRUE),
  .after = week_ending
) 


covid_full %>% head()
```
 
 
 
-----gender------

looking into time this time 
`specialty is NA ?????? `
```{r}
covid_sex <- covid_full %>% 
  filter(admission_type == "All",
         age_group == "All ages",
         sex != "All") %>% 
  group_by(sex,week_ending) %>% 
  summarise(total_admissions = sum(number_admissions))

covid_sex
```


```{r}
covid_sex %>% ggplot(aes(week_ending, total_admissions))+
  geom_line(aes(col = sex))+
  labs(title = "Female/male over time ")
```
```{r}
 covid_full %>% 
  filter(admission_type == "All",
         age_group == "All ages",
         sex != "All") %>% 
  group_by(sex,month) %>% 
  summarise(total_admissions = sum(number_admissions))%>% 
  ggplot(aes(month, total_admissions))+
  geom_line(aes(col = sex, group = sex))+
  geom_point(aes(col = sex))+
  labs(title = "Female/male over time (month)")
```
 
 
 
 
```{r}
covid_simd<- covid_full %>% 
  drop_na(simd_quintile) %>% 
  group_by(simd_quintile,week_ending) %>% 
  summarise(total_admissions = sum(number_admissions))

covid_simd
```
```{r}
covid_simd %>% ggplot(aes(week_ending, total_admissions))+
  geom_line(aes(col = as.factor(simd_quintile)))+
  labs(title = "simd over time")
```

```{r}
covid_simd %>% ggplot(aes(simd_quintile, total_admissions))+
  geom_col(aes(fill = as.factor(simd_quintile)))+
  labs(title = "simd compare to each other")
```
 sex-test --------
 female = male
 female > male 
 alpha = 0.05
```{r}
covid_sex %>% 
  ggplot(aes(sex, total_admissions)) + 
  geom_boxplot(aes(col = sex))+
  labs(title= "female, male title admissions")
```
```{r}

null_distribution <- covid_sex %>% 
  specify(total_admissions ~ sex) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("Female", "Male")) 


observed_stat <- covid_sex %>% 
  specify(total_admissions ~ sex) %>%
  calculate(stat = "diff in means", order = c("Female", "Male"))
  
observed_stat


null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "right")

p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "right")

p_value
```
 

```{r}

specialties_raw <- read_csv(here("raw_data/non_covid/inpatient_and_daycase_by_nhs_board_of_treatment_and_specialty.csv")) %>%
 janitor:: clean_names()


```

```{r}

specialties <- specialties_raw %>% 
  inner_join(hb, "hb") %>% 
  select(-ends_with("qf"), -hb, - specialty) %>% 
  select(quarter, hb_name, everything()) %>% 
  mutate(year = as.numeric(str_sub(quarter,1, 4)), .before = quarter) %>% 
  mutate(is_covid_year = case_when(
    year >= 2020 ~ TRUE,
    year < 2020 ~ FALSE
  ))

specialties <- specialties %>% mutate(hb_name = str_sub(hb_name, start = 5)) %>% 
  select(is_covid_year,hb_name,admission_type,episodes, specialty_name)

write_csv(specialties, here("clean_data/specialties.csv"))
```

```{r}

total_admissions <- specialties %>% 
  #filter(hb_name %in% c(	"NHS Ayrshire and Arran","NHS Dumfries and Galloway")) %>% 
  group_by(is_covid_year) %>% 
  filter(admission_type == "All Inpatients and Day cases") %>% 
  summarise(total_admissions = sum(episodes)) %>% pull()
  


change_in_specialties <- specialties %>% 
 # filter(hb_name %in% c("NHS Ayrshire and Arran","NHS Dumfries and Galloway")) %>% 
  group_by(specialty_name, is_covid_year) %>% 
  summarise(total_episodes = sum(episodes)) %>% 
  pivot_wider(names_from = is_covid_year, values_from = total_episodes) %>% 
  rename("covid_year" = "TRUE", "pre_covid_year" = "FALSE") %>% 
  mutate(pre_covid_year_prop = pre_covid_year / total_admissions[1],
         covid_year_prop = covid_year /total_admissions[2],
         percentage_change = covid_year_prop - pre_covid_year_prop)

 

change_in_specialties %>% 
  arrange(desc(percentage_change)) %>% 
  head(5) %>% 
  filter(percentage_change > 0) %>% 
  ggplot(aes(x = reorder(specialty_name, percentage_change, decreasing = TRUE),
             y = percentage_change)) +
  geom_col(aes(fill = specialty_name)) +
  theme_classic() +
  scale_y_continuous(labels = percent_format())+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8,angle = 45, hjust = 1)) +
  labs(y = "Percentage Increase (%)",
       title = "Increase in hospital admissions (by specialty) - pre-Covid vs Covid",
       subtitle = "More than 1,000 admissions")

```
```{r}
change_in_specialties %>% 
  arrange(desc(abs(percentage_change)))%>% 
  head(5) %>% 
  ggplot(aes(x = reorder(specialty_name, percentage_change, decreasing = TRUE),
             y = percentage_change)) +
  geom_col(aes(fill = percentage_change >= 0 )) +
  theme_classic() +
  scale_y_continuous(labels = percent_format())+
  scale_fill_manual(values = c("FALSE" = "seagreen", "TRUE"= "red"))+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8,angle = 45, hjust = 1)) +
  labs(y = "Percentage Increase (%)",
       title = "Top 5 changes in hospital admissions (by specialty) - pre-Covid vs Covid",
       subtitle = "+tive and -tive changes")

write.csv(change_in_specialties, here("clean_data/specialties.csv"))

```



maybe useful 

library(reshape2)
before_after_covid <- specialties %>% 
    filter(hb_name == "NHS Ayrshire and Arran") %>% #for input 
  group_by(specialty_name, is_covid_year,hb_name) %>% 
  summarise(total_episodes = sum(episodes)) %>% 
  pivot_wider(names_from = is_covid_year, values_from = total_episodes) %>% 
  rename("covid_year" = "TRUE", "pre_covid_year" = "FALSE") %>% 
  mutate(pre_covid_year_prop = pre_covid_year / total_admissions[1],
         covid_year_prop = covid_year / total_admissions[2],
         percentage_change = ((covid_year_prop / pre_covid_year_prop) - 1) * 100) %>% 
  select(specialty_name,hb_name,pre_covid_year_prop,covid_year_prop,
percentage_change
)
  

data_mod_before_after_covid <- melt(before_after_covid,
                  measure.vars=c('pre_covid_year_prop', 'covid_year_prop')) %>% 
  arrange(desc(percentage_change)) %>% 
  filter(percentage_change > 0) %>% 
  head(10)
  
  

ggplot(data_mod_before_after_covid) +
geom_col(aes(x = reorder(specialty_name, percentage_change, decreasing = TRUE),
             y=value, color=variable, fill = variable), position = "dodge")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(size = 8,angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("pre_covid_year_prop" = "grey", "covid_year_prop"= "red"))+
  scale_colour_manual(values = c("pre_covid_year_prop" = "grey", "covid_year_prop"= "red"))+
  labs(y = "Percentage (%)",
       title = "Top 5 changes in hospital admissions (by specialty) - pre-Covid vs Covid",
       subtitle = "y is the percentage but x is ordered by percentage changes")
  
  

```{r}
 total_admissions <- specialties %>% 
    #filter(hb_name %in% input$health_board_input) %>% 
    group_by(is_covid_year) %>% 
    filter(admission_type == "All Inpatients and Day cases") 
    summarise(total_admissions = sum(episodes)) %>% full_join(specialties)%>% 
    #filter(hb_name %in% input$health_board_input) %>% 
    group_by(specialty_name, is_covid_year) %>% 
    summarise(total_episodes = sum(episodes)) %>% 
    pivot_wider(names_from = is_covid_year, values_from = total_episodes) %>% 
    rename("covid_year" = "TRUE", "pre_covid_year" = "FALSE") %>% 
    mutate(pre_covid_year_prop = pre_covid_year / total_admissions[1],
           covid_year_prop = covid_year /total_admissions[2],
           percentage_change = covid_year_prop - pre_covid_year_prop)%>% 
      arrange(desc(percentage_change)) %>% head(5) %>% 
      filter(percentage_change > 0)  

```






 