---
title: "R Notebook"
output: html_notebook
---
# loading in relevant packages and data
```{r}
library(tidyverse)
library(janitor)
library(plotly)
library(lubridate)
hospital_admissions_by_hb_agesex_2022_03_02 <- read_csv("../raw_data/covid_data/hospital_admissions_hb_agesex_20220302.csv")# %>% view()
hospital_admissions_by_hb_simd_2022_03_02 <- read_csv("../raw_data/covid_data/hospital_admissions_hb_simd_20220302.csv")# %>% view()
hospital_admissions_by_hb_speciality_2022_03_02 <- read_csv("../raw_data/covid_data/hospital_admissions_hb_specialty_20220302.csv")# %>% view()
# hospital_admissions_by_hscp_agesex_2022_03_02 <- read_csv("covid_data/hospital_admissions_hscp_agesex_20220302.csv")# %>% view()
# hospital_admissions_by_hscp_simd_2022_03_02 <- read_csv("covid_data/hospital_admissions_hscp_simd_20220302.csv")# %>% view()
# hospital_admissions_by_hscp_speciality_2022_03_02 <- read_csv("covid_data/hospital_admissions_hscp_specialty_20220302.csv")# %>% view()
inpatient_and_daycase_by_nhs_board_of_treatment_age_and_sex <- read_csv("../raw_data/non_covid/inpatient_and_daycase_by_nhs_board_of_treatment_age_and_sex.csv")

#view(hospital_admissions_by_hb_agesex_2022_03_02)
#view(hospital_admissions_by_hb_simd_2022_03_02)
#view(hospital_admissions_by_hb_speciality_2022_03_02)
#view(inpatient_and_daycase_by_nhs_board_of_treatment_age_and_sex)
```

# Data integrity - I like to see if / where there are any missing values.
#### We can see that the missing data is distributed to those columns relating to Qualified Fields eg HBQF, AgeGroup and SexQF
#### Initial thoughts, we need to clean the data with regards to:
#### ~ names
#### ~ dates
```{r}
hospital_admissions_by_hb_agesex_2022_03_02 %>% 
  distinct(WeekEnding, HB, AgeGroup, Sex, .keep_all = TRUE) %>% 
  summarise(across(.fns = ~ sum(is.na(.x))))# %>% 
#  ggplot()+
#  geom_histogram(aes(x = HBQF))

```
## Making relevant amendments based on my obeservation (lines 24 to 25) and other requirements detailed below:
### Amending data or Removing columns based on their value / usefulness to answering our business objectives eg
### Formatting the date and creating a new column called 'week_ending_amended'
### Factoring the age_group column to a particular order and putting it in a column called age_group_fct
### ~ removing the following columns: hbqf, age_group_qf, sex_qf, admission_type_qf
```{r}
# cleaning names
hospital_admissions_by_hb_agesex_2022_03_02_cleaned <- hospital_admissions_by_hb_agesex_2022_03_02 %>% 
  clean_names() %>%
  mutate(week_ending_amended = as.Date(ymd(week_ending)), .after = week_ending) %>% 
  mutate(month_of_admission = month(week_ending_amended, label = T), .before = hb) %>% 
  mutate(year_of_admission = year(week_ending_amended), .before = hb) %>% 
  mutate(month_year = make_datetime(year_of_admission, month_of_admission), .before = hb) %>%
  mutate(age_group_fct = factor(age_group, levels = c("Under 5", "5 - 14", "15 - 44", "45 - 64", "65 - 74", "75 - 84", "85 and over", "All ages")), .after = age_group) %>% 
  select(week_ending, week_ending_amended, month_of_admission, year_of_admission, month_year, hb, age_group, age_group_fct, sex, admission_type, number_admissions, average20182019, percent_variation) # removing columns that are not required

hospital_admissions_by_hb_agesex_2022_03_02_cleaned

# to be removed
# hospital_admissions_by_hb_agesex_2022_03_02_cleaned_names_columns_removed <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>% 
#  select(week_ending, week_ending_amended, hb, age_group, age_group_fct, sex, admission_type, number_admissions, average20182019, # percent_variation)
# hospital_admissions_by_hb_agesex_2022_03_02_cleaned_names_columns_removed
```

# For plotly only
### ~ (plotly only) removing admission_type = All from full dataset and 
### ~ (plotly only) for age removing sex = Female | Male
### ~ (plotly only) for sex removing age = All ages
```{r}
## Removing rows and columns based on their value / usefulness to answering our business objectives eg
### ~ removing the following columns: hbqf, age_group_qf, sex_qf, admission_type_qf
### ~ removing admission_type = All from full dataset and 
### ~ for age removing sex = Female | Male
### ~ for sex removing age = All ages

# this removes all the 'admission_type' rows where the value is "All"
 row_removal_value_admission_type_All_plotly <- subset(hospital_admissions_by_hb_agesex_2022_03_02_cleaned, admission_type != "All")
 row_removal_value_admission_type_All_plotly

# this is specific for 'age_group' only and removes all the 'age_group' rows where the value is "All ages"
 row_removal_value_admission_type_and_age_group_All <- subset(row_removal_value_admission_type_All_plotly, age_group != "All ages",age_group_fct != "All ages")
 row_removal_value_admission_type_and_age_group_All

# this is specific for 'sex' only and removes all the 'sex' rows where the value is "All"
 row_removal_value_admission_type_and_sex_All <- subset(row_removal_value_admission_type_All_plotly, sex != "All")
 row_removal_value_admission_type_and_sex_All
```


# By age
```{r}
#hospital_admissions_by_hb_agesex_test_plotly <- plot_ly(data = row_removal_value_admission_type_and_age_group_All,
#                      x = ~number_admissions,
#                      y = ~week_ending_amended)
#hospital_admissions_by_hb_agesex_test_plotly


# FOR SHINY APP

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(hb) %>%
  summarise(week_ending_amended, age_group_fct, hb, number_admissions) %>% 
  distinct()


# by age vs number of admissions
hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(hb) %>%
 # ungroup() %>% 
  summarise(week_ending_amended, age_group_fct, hb, number_admissions) %>% 
  distinct() %>% 
  ggplot()+
  aes(x = age_group_fct, y = number_admissions)+
  geom_col()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

# by week_ending_amended vs number of admissions split by age group

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type != "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(hb) %>%
 # ungroup() %>% 
  summarise(week_ending_amended, age_group_fct, hb, number_admissions) %>% 
  distinct() 

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type != "All") %>% 
  filter(age_group_fct != "All ages") %>%
  filter(sex == "All") %>% 
  distinct(age_group_fct, hb, month_year, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, month_year, number_admissions) %>% 
  group_by(hb) %>%
  #ungroup() %>% 
  summarise(month_year, age_group_fct, hb, number_admissions) %>% 
  ggplot()+
  aes(x = month_year, y = number_admissions, fill = age_group_fct)+
  geom_col()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

# by week_ending_amended vs number of admissions split by age group
  total_adm_equal <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  filter(sex == "All") %>%   
  select(hb, month_year, month_of_admission, number_admissions, age_group_fct) %>% 
  distinct() %>% 
  group_by(month_year, age_group_fct) %>%
  #distinct(month_of_admission, number_admissions, .keep_all = TRUE) %>% 
  summarise(sum_admissions = sum(number_admissions)) %>% 
#  distinct()
  ggplot()+
  aes(x = month_year, y = sum_admissions, group = age_group_fct, col = age_group_fct)+
  geom_line()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

  total_adm_equal
```

# By sex
```{r}
hospital_admissions_by_hb_test_agesex <- plot_ly(data = row_removal_value_admission_type_and_sex_All,
                      x=~week_ending_amended,
                      y=~sex)
hospital_admissions_by_hb_test_agesex

# by sex -table
hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  distinct(week_ending_amended, hb, sex, number_admissions, admission_type, .keep_all = TRUE) %>% 
#  filter(hb == "S08000015") %>% 
#  filter(sex == "Male") %>% 
#  filter(week_ending_amended == "2020-01-12") %>% 
  select(week_ending_amended, hb, sex, number_admissions, admission_type) %>% 
  group_by(hb) %>% 
  count(week_ending_amended, sex, number_admissions)

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  distinct(week_ending_amended, hb, sex, number_admissions, admission_type, .keep_all = TRUE) %>% 
#  filter(hb == "S08000015") %>% 
#  filter(sex == "Male") %>% 
#  filter(week_ending_amended == "2020-01-12") %>%
  select(week_ending_amended, hb, sex, number_admissions, admission_type) %>% 
  group_by(hb) %>% 
  count(week_ending_amended, sex, number_admissions) %>% 
  ggplot()+
  aes(x = sex, y = number_admissions)+
  geom_col()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  distinct(week_ending_amended, hb, sex, number_admissions, admission_type, .keep_all = TRUE) %>% 
  filter(admission_type != "All") %>% 
  filter(sex != "All") %>% 
  select(week_ending_amended, hb, sex, number_admissions, admission_type) %>% 
  group_by(hb) %>% 
  summarise(week_ending_amended, sex, number_admissions)

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  distinct(week_ending_amended, hb, sex, number_admissions, admission_type, .keep_all = TRUE) %>% 
  filter(admission_type != "All") %>% 
  filter(sex != "All") %>% 
  select(week_ending_amended, hb, sex, number_admissions, admission_type) %>% 
  group_by(hb) %>% 
  summarise(week_ending_amended, sex, number_admissions) %>% 
  ggplot()+
  aes(x = sex, y = number_admissions)+
  geom_col()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")
```

# test section
```{r}
# work out proportional split
  age_group_prop <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(sex == "All") %>% 
  filter(admission_type == "All") %>% 
  pivot_wider(names_from = "age_group_fct", values_from = "age_group_fct") %>%
  clean_names() %>% 
  mutate(under_5 = as.character(under_5),
         x5_14   = as.character(x5_14),
         x15_44  = as.character(x15_44),
         x45_64  = as.character(x45_64),
         x65_74  = as.character(x65_74),
         x75_84  = as.character(x75_84),
        x85_and_over= as.character(x85_and_over),
        all_ages = as.character(all_ages)) %>% 
  mutate(under_5 = coalesce(under_5, "0"),
         x5_14 = coalesce(x5_14, "0"),
         x15_44 = coalesce(x15_44, "0"),
         x45_64 = coalesce(x45_64, "0"),
         x65_74 = coalesce(x65_74, "0"),
         x75_84 = coalesce(x75_84, "0"),
         x85_and_over = coalesce(x85_and_over, "0"),
         all_ages = coalesce(all_ages, "0")) %>% 
  mutate(under_5 = as.numeric(recode(under_5, "Under 5" = "1")),
         x5_14 = as.numeric(recode(x5_14, "5 - 14" = "1")),
         x15_44 = as.numeric(recode(x15_44, "15 - 44" = "1")),
         x45_64 = as.numeric(recode(x45_64, "45 - 64" = "1")),
         x65_74 = as.numeric(recode(x65_74, "65 - 74" = "1")),
         x75_84 = as.numeric(recode(x75_84, "75 - 84" = "1")),
         x85_and_over = as.numeric(recode(x85_and_over, "85 and over" = "1")))

age_group_prop_test <- subset(age_group_prop, all_ages != "All ages") # removes rows where all_ages contains "All ages"

sum_age_group_prop_test <- age_group_prop_test %>% 
summarise(under_5_prop = sum(under_5)/nrow(age_group_prop),
          x5_14_prop = sum(x5_14)/nrow(age_group_prop),
          x15_44_prop = sum(x15_44)/nrow(age_group_prop),
          x45_64_prop = sum(x45_64)/nrow(age_group_prop),
          x65_74_prop = sum(x65_74)/nrow(age_group_prop),
          x75_84_prop = sum(x75_84)/nrow(age_group_prop),
          x85_and_over = sum(x85_and_over)/nrow(age_group_prop))
sum_age_group_prop_test

quick_test <- age_group_prop %>% 
  mutate(ages_prop = as.numeric(if_else(under_5 == "1", number_admissions*sum_age_group_prop_test$under_5_prop, 0)),
         ages_prop = as.numeric(if_else(x5_14 == "1", number_admissions*sum_age_group_prop_test$x5_14_prop, 0)),
         x15_44 = as.numeric(recode(x15_44, "1" = number_admissions*sum_age_group_prop_test$x15_44_prop)),
         x45_64 = as.numeric(recode(x45_64, "1" = number_admissions*sum_age_group_prop_test$x45_64_prop)),
         x65_74 = as.numeric(recode(x65_74, "1" = number_admissions*sum_age_group_prop_test$x65_74_prop)),
         x75_84 = as.numeric(recode(x75_84, "1" = number_admissions*sum_age_group_prop_test$x75_84_prop)),
         x85_and_over = as.numeric(recode(x85_and_over, "1" = number_admissions*sum_age_group_prop_test$x85_and_over)))
quick_test %>% 
  ggplot()+
  aes(x = month_year, y = sum_admissions, group = age_group_fct, col = age_group_fct)+
  geom_line()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")


  age_group_prop_test <- subset(age_group_prop, all_ages == "All ages")
  
    summarise(age_group_fct_prop = count(age_group_fct) / nrow(hospital_admissions_by_hb_agesex_2022_03_02_cleaned))

  age_group_prop <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>% 
        mutate(age_group_prop = age_group_fct) %>% 
    mutate(age_group_prop = age_group_fct) %>% 
    filter(age_group_prop != "All ages") %>% 
    pivot_longer(contains("Under 5"), age_group_prop, names_to = "test", values_to = "test2")
count(prop = age_group_prop / n)

```



```{r}
month_test <- hospital_admissions_by_hb_agesex_2022_03_02 %>%
  mutate(week_ending_amended = as.Date(ymd(WeekEnding)), .after = WeekEnding) %>% 
  mutate(month_amended = month(week_ending_amended, label = T))

month_test <- month_test %>% 
    mutate(week_ending_amended = mdy(week_ending_amended)) %>% 

month_test
```

```{r}
month_test <- hospital_admissions_by_hb_agesex_2022_03_02 %>% 
  mutate(week_ending_amended = as.character(ymd(WeekEnding)), .after = WeekEnding) %>% 
  mutate(month_amended = as.numeric(month(week_ending_amended, label = T)))
month_test

row_removal_value_admission_type_and_age_group_All %>%
  #filter(hb == "S08000015") %>% 
  #filter(age_group_fct == "Under 5") %>% 
  #filter(week_ending_amended == "2020-01-12") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(hb) %>%
  #ungroup() %>% 
  summarise(week_ending_amended, age_group_fct, hb, number_admissions)

# by age vs number of admissions
row_removal_value_admission_type_and_age_group_All %>%
  #filter(hb == "S08000015") %>% 
  #filter(age_group_fct == "Under 5") %>% 
  #filter(week_ending_amended == "2020-01-12") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(hb) %>%
  #ungroup() %>% 
  summarise(week_ending_amended, age_group_fct, hb, number_admissions) %>% 
  ggplot()+
  aes(x = age_group_fct, y = number_admissions)+
  geom_col()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(week_ending_amended) %>%
  #ungroup() %>% 
  summarise(week_ending_amended,
            age_group_fct,
            number_of_admissions = sum(number_admissions)) %>% 
  distinct() %>% 
  ggplot()+
  aes(x = week_ending_amended, y = number_of_admissions, fill = age_group_fct)+
  geom_col()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  distinct(age_group_fct, hb, week_ending_amended, number_admissions, .keep_all = TRUE) %>% 
  select(age_group_fct, hb, week_ending_amended, number_admissions) %>% 
  group_by(week_ending_amended) %>%
  #ungroup() %>% 
  summarise(week_ending_amended,
            age_group_fct,
            number_of_admissions = sum(number_admissions)) %>% 
  distinct()

 total_adm_notequal <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type != "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  distinct(week_ending_amended, hb, number_admissions, .keep_all = TRUE) %>% 
  select(number_admissions) %>% 
#  group_by(number_admissions) %>%
  #ungroup() %>% 
  summarise(number_of_admissions = sum(number_admissions)) #%>% 
#  distinct()

  total_adm_equal <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  filter(sex == "All") %>%   
  select(hb, week_ending_amended, month_of_admission, number_admissions, age_group_fct) %>% 
  distinct() %>% 
  group_by(month_of_admission, age_group_fct) %>%
  #distinct(month_of_admission, number_admissions, .keep_all = TRUE) %>% 
  summarise(sum_admissions = sum(number_admissions)) %>% 
#  distinct()
  ggplot()+
  aes(x = month_of_admission, y = sum_admissions, group = age_group_fct, col = age_group_fct)+
  geom_line()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

  total_adm_equal
  total_adm_notequal
  
  total_adm_equal_prop <- hospital_admissions_by_hb_agesex_2022_03_02_cleaned %>%
  filter(admission_type == "All") %>% 
  filter(age_group_fct != "All ages") %>% 
  filter(sex == "All") %>%   
  select(hb, week_ending_amended, month_of_admission, number_admissions, age_group_fct) %>% 
  distinct() %>% 
  group_by(month_of_admission, age_group_fct) %>%
  #distinct(month_of_admission, number_admissions, .keep_all = TRUE) %>% 
  summarise(sum_age_group_fct = count(age_group_fct)) %>% 
#  distinct()
  ggplot()+
  aes(x = month_of_admission, y = sum_admissions, group = age_group_fct, col = age_group_fct)+
  geom_line()+
  xlab("tbc")+
  ylab("tbc")+
  ggtitle("tbc")

  total_adm_equal
```

