---
title: "COVID Datasets"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)

```

# load in health board data
```{r}

hb <- read_csv(here("raw_data/general/hb14_hb19.csv")) %>% 
  clean_names() %>% 
  select(hb, hb_name)

```


# Demographic data - COVID

```{r}

admissions_raw <- read_csv(here("raw_data/covid_data/hospital_admissions_hb_agesex_20220302.csv")) %>% 
  clean_names() %>% 
  mutate(week_ending = ymd(week_ending))

```

# clean admissions data (age & sex)
```{r}

admissions <- admissions_raw %>% 
  left_join(hb, "hb") %>% 
  select(week_ending, hb_name, age_group, sex, admission_type, number_admissions, average20182019, percent_variation)

```
## Analysis

## Total admissions (all ages and sexes, all departments)
```{r}

admissions %>% 
  filter(age_group == "All ages" & sex == "All") %>% 
  group_by(week_ending) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions)) +
  geom_line() +
  labs(title = "Total Admissions") +
 theme_classic() +
  labs(y = "Total Admissions",
       col = "Age Group") +
  theme(axis.title.x = element_blank())

```

## Total admissions (by sex, all ages/departments)
```{r}

admissions %>% 
  filter(age_group == "All ages" & sex != "All" & admission_type == "All") %>% 
  group_by(week_ending, sex) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions, colour = sex)) +
  geom_line() +
  labs(title = "Total Admissions - split by sex") +
  theme_classic()
#2.7. Interpreting the results of statistical analysis
# number of males and females are pretty much the same across the entire covid timeline

```

```{r}

total_all_sexes <- admissions %>% 
  filter(age_group == "All ages" & sex == "All" & admission_type == "All") %>% 
  summarise(total_admissions = sum(number_admissions))

total_m_and_f <- admissions %>% 
  filter(age_group == "All ages" & sex != "All" & admission_type == "All") %>% 
  summarise(total_admissions = sum(number_admissions))

diff <- total_all_sexes - total_m_and_f
diff

fraction <- diff / total_all_sexes * 100
fraction

```


## Total admissions (by age, all sexes/departments)
```{r}

admissions %>%
  mutate(age_group = factor(age_group, levels = c("Under 5",
                                                  "5 - 14",
                                                  "15 - 44",
                                                  "45 - 64",
                                                  "65 - 74",
                                                  "75 - 84",
                                                  "85 and over"
                                                  ))) %>% 
  filter(age_group != "All ages" & sex == "All" & admission_type == "All") %>% 
  group_by(week_ending, age_group) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions, colour = age_group)) +
  geom_line() +
  labs(title = "Total Admissions - split by age group") +
  theme_classic() +
  labs(y = "Total Admissions",
       col = "Age Group") +
  theme(axis.title.x = element_blank())

```


```{r}
## Total admissions (by admission type, all sexes/age_groups)

admissions %>% 
  distinct(admission_type)


admissions %>% 
  filter(age_group == "All ages" & sex == "All" & admission_type != "All") %>% 
  group_by(week_ending, admission_type) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions, colour = admission_type)) +
  geom_line() +
  labs(title = "Total Admissions - split by admission type") +
  theme_classic() +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Total Admissions",
       col = "Admission Type") +
  theme(axis.title.x = element_blank(),
        legend.title = element_text(face = "bold"))

```

```{r}

specialty_raw <- read_csv(here("raw_data/covid_data/hospital_admissions_hb_specialty_20220302.csv")) %>% 
  clean_names() %>% 
  mutate(week_ending = ymd(week_ending))

```

## Clean data (specialty)
```{r}

specialty <- specialty_raw %>% 
  left_join(hb, "hb") %>% 
  select(week_ending, hb_name, admission_type, specialty, number_admissions, average20182019, percent_variation)

```

### Analyse data (specialty)

### the number of total admissions?

```{r}

specialty %>% 
  filter(admission_type == "All" & specialty == "All") %>% 
  group_by(week_ending) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions)) +
  geom_line()

# looks like the same as the previous data

```

### What is the number of admissions per admission type?

```{r}

specialty %>% 
  filter(admission_type != "All" & specialty == "All") %>% 
  group_by(week_ending, admission_type) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions, col = admission_type)) +
  geom_line() +
  theme_classic()


# looks the same as the previous plot

```

### What is the number of admissions per specialty?

```{r}
# list of unique admission numbers, i.e. the list of specialties that add up to the total admissions value 
specialty_list <- c("Accident & Emergency", "Community", "Gynaecology", "Medical (incl. Cardiology & Cancer)", "Paediatrics (medical & surgical)", "Surgery")

specialty %>% 
  filter(admission_type == "All" & specialty != "All") %>% 
  group_by(specialty) %>% 
  summarise(mean_admissions = mean(number_admissions)) %>% 
  arrange(desc(mean_admissions))

specialty %>% 
  filter(admission_type == "All" & specialty %in% specialty_list) %>% 
  group_by(week_ending, specialty) %>% 
  summarise(total_admissions = sum(number_admissions)) %>% 
  ggplot(aes(x = week_ending, y = total_admissions, col = specialty)) +
  geom_line() +
  theme_classic() +
  # facet_wrap(~specialty, scales = "free_y") +
  theme(legend.position = "bottom",
        axis.title.x = element_blank()) +
  labs(y = "Total Admissions",
       title = "Breakdown of total admissions per specialty")

# the top specialties are Medical (incl. cardiology & cancer) and surgery.

```
